Data preparation: Study 1

### 2021-07-08 paper_study1.Rmd
ajustes
- cambio a data_s1.RData
- Modificación en data prep de study1. Se detalla mejor el procedimiento.
- check render desde consola  html (OK) y pdf (OK)

```{r eval=FALSE, include=FALSE}
# for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
# clean #in the yml
rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::html_document2")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.pos = "H",
	message = FALSE,
	warning = FALSE,
	cache = FALSE
)
# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
options(knitr.graphics.error = FALSE)
Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

@@

@@

knitr::kable(tableissp, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
  kableExtra::kable_styling(
    full_width = F,
    latex_options = c("hold_position"),
    position = "center",
    font_size = 20,
    bootstrap_options=c("striped", "bordered")) %>%
  kableExtra::column_spec(column = 1, width = "2cm", ) %>%
  kableExtra::column_spec(column = 2,width = "5cm") %>%
  kableExtra::column_spec(column = 3,width = "12cm") %>%
  kableExtra::collapse_rows(columns = 1:2,valign = "middle")
```

@@

```{r desc-issp, eval=FALSE, include=FALSE, results='asis'}
load(file = "input/data/proc/merit_per_pref.rda")
pacman::p_load(dplyr)
table_format = if(knitr::is_html_output()) {
  "html"
} else if(knitr::is_latex_output()) {
  "latex"
}

@@

merit_per_pref %>% 
  dplyr::select(hwork,ambition,wealthy,pareduc,
                race,gender,people,polcone,welljob,hardjob,family,child) %>% 
  psych::describe() %>% 
  as.data.frame() %>% 
  dplyr::select("Mean"=mean,"SD"=sd,"Min"=min,"Max"=max) %>% 
  round(2) ->desc.issp

@@

```{r}
load(file = "input/data/proc/merit_per_pref.rda")
pacman::p_load(dplyr)
likert_issp<- merit_per_pref %>% 
  dplyr::select(hwork,ambition,wealthy,pareduc,
                race,gender,people,polcone,
                welljob,hardjob,family,child) %>% na.omit()

@@

```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.height='80%'}
cowplot::add_sub(
  cowplot::plot_grid(p1,p2,p3,p4,leg,
    align = "hv",
    axis = "a",
    nrow = 5,
    ncol = 1,
    rel_heights = c(1,2,1,1,0.25)
    ),
  label = paste0(
    "Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
    " (n=",dim(likert_issp)[1],")"
    ),
  size = 10,
  hjust = 0.25
  ) %>% 
  cowplot::ggdraw() 
```

@@

```{r corr-issp,fig.cap = "Perceptions & preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
pacman::p_load(dplyr)
cor_issp<- 
  merit_per_pref %>%
  dplyr::select(hwork,ambition,wealthy,pareduc,
                race,gender,people,polcone,
                welljob,hardjob,family,child) %>% 
  lavaan::lavCor(., ordered=names(.))
diag(cor_issp) = NA
rownames(cor_issp) <- c("A. Hard work",
                       "B. Having ambition",
                       "C. Wealthy family",
                       "D. Educated parents",
                       "E. Race",
                       "F. Gender",
                       "G. Knowing people",
                       "H. Political connections",
                       "I. Well job",
                       "J. Hard job",
                       "K. Support family",
                       "L. Has children")
colnames(cor_issp) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", 
                       "(H)", "(I)","(J)","(K)","(L)")
corrplot::corrplot(cor_issp,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(12),
  bg = "white",
  na.label = "-",
  lower.col = "black")
```

@@

```{r sum-fit-issp}
load(file = "input/data/proc/merit_per_pref.rda")
pacman::p_load(dplyr, kableExtra, knitr)

cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,ordered = T)

cfa_perpref2 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  parents =~ wealthy + pareduc
	  backgrd =~ race + gender
	  networks=~ people + polcone
	  percnmerit=~ parents + backgrd + networks
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = merit_per_pref,ordered = T)

@@

# Create table
table_issp_fits <-
  knitr::kable(sum_fit_issp, format="html", digits=3, booktabs=T, 
               col.names=colnames,  
               caption = "Summary fit indices according to model", 
               escape = FALSE) %>%
  kableExtra::kable_styling(full_width = F, 
                            font_size = 15,
                            bootstrap_options=c("striped", "bordered"))
table_issp_fits
```

Table \@ref(tab:sum-fit-issp) shows the four-factor model fit allowing comparison of first and second-order factors. The model restricted to four first-order factor dimensions shows a regular fit (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)) in contrast to the model considering a second-order factor for non-meritocratic perception that has a slightly better fit based on the standard fit measures (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535).

@@

```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy"}
knitr::include_graphics('output/images/cfa_issp_study1.PNG')
```

@@

```{r tb-perpref1, echo=FALSE}
load(file = "input/data/proc/merit_per_pref.rda")
pacman::p_load(dplyr,knitr,kableExtra)
options(knitr.kable.NA = '')
table_format = if(knitr::is_html_output()) {
  "html"
} else if(knitr::is_latex_output()) {
  "latex"
}

@@

cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
                            ordered = c("wealthy","pareduc","ambition","hwork",
                                        "people","polcone","race","gender",
                                        "respons","yeduc",
                                        "family","child",
                                        "welljob","hardjob"))
labs <- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Reasons for pay: well job",
  "Reasons for pay: hard job",
  "Reasons for pay: support family",
  "Reasons for pay: has children"
)

@@

# Table --------------------------------------------------------------------.
tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA = Root mean square error of approximation")
tb.col <- c("","Meritocratic","Non-meritocratic",
            "Meritocratic","Non-meritocratic")
tb.caption <- c(" Factor loadings and fit measures for ISSP 2009")
tb.issp<- kable(tb.perpref1,escape = FALSE,digits = 3,
                align = "lcccc",
                col.names = tb.col, 
                caption = tb.caption, 
                booktabs = T,
                linesep = "",
                format = table_format) %>%
          kable_styling(full_width = table_format2,
                        latex_options = c("hold_position"), 
                        position = "center",
                        font_size = 12) %>%
          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
          row_spec(row = 12,hline_after = TRUE) %>%
          add_indent(c(13:17)) %>%
          footnote(general =tb.foot ,footnote_as_chunk = T);tb.issp
```

>>>>

```{r eval=FALSE, include=FALSE}
rmarkdown::render("paper_study1.Rmd", output_format = "bookdown::pdf_document2")
rmarkdown::render("paper_study1.Rmd", output_format = "bookdown::html_document2")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.pos = "H",
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	cache = FALSE, 
	out.width = if(knitr::is_html_output()) {"80%"} 
	            else if(knitr::is_latex_output()) {"80%"}
)
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
options(knitr.graphics.error = FALSE)
Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

@@

fsize = if(knitr::is_html_output()) {
  20
} else if(knitr::is_latex_output()) {
  12
}

@@

knitr::kable(tableissp, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
  kableExtra::kable_styling(
    full_width = F,
    latex_options = c("hold_position"),
    position = "center",
    font_size = fsize,
    bootstrap_options=c("striped", "bordered")) %>%
  kableExtra::column_spec(column = 1, width = "2cm", ) %>%
  kableExtra::column_spec(column = 2,width = "5cm") %>%
  # kableExtra::column_spec(column = 3,width = "12cm") %>%
  kableExtra::collapse_rows(columns = 1:2,valign = "middle")
```

@@

```{r desc-issp, eval=FALSE, include=FALSE, results='asis'}
load(file = "input/data/proc/data_s1.RData")
pacman::p_load(dplyr)
table_format = if(knitr::is_html_output()) {
  "html"
} else if(knitr::is_latex_output()) {
  "latex"
}

@@

data_s1 %>% 
  dplyr::select(hwork,ambition,wealthy,pareduc,
                race,gender,people,polcone,welljob,hardjob,family,child) %>% 
  psych::describe() %>% 
  as.data.frame() %>% 
  dplyr::select("Mean"=mean,"SD"=sd,"Min"=min,"Max"=max) %>% 
  round(2) ->desc.issp

@@

```{r}
load(file = "input/data/proc/data_s1.RData")
pacman::p_load(dplyr)
likert_issp<- data_s1 %>% 
  dplyr::select(hwork,ambition,wealthy,pareduc,
                race,gender,people,polcone,
                welljob,hardjob,family,child) %>% na.omit()

@@

```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items"}
cowplot::add_sub(
  cowplot::plot_grid(p1,p2,p3,p4,leg,
    align = "hv",
    axis = "a",
    nrow = 5,
    ncol = 1,
    rel_heights = c(1,2,1,1,0.25)
    ),
  label = paste0(
    "Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
    " (n=",dim(likert_issp)[1],")"
    ),
  size = 10,
  hjust = 0.25
  ) %>% 
  cowplot::ggdraw() 
```

@@

```{r corr-issp,fig.cap = "Perceptions and preferences for ISSP meritocracy items polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
load(file = "input/data/proc/data_s1.RData")
pacman::p_load(dplyr)
cor_issp<- 
  data_s1 %>%
  dplyr::select(hwork,ambition,wealthy,pareduc,
                race,gender,people,polcone,
                welljob,hardjob,family,child) %>% 
  lavaan::lavCor(., ordered=names(.))
diag(cor_issp) = NA
rownames(cor_issp) <- c("A. Hard work",
                       "B. Having ambition",
                       "C. Wealthy family",
                       "D. Educated parents",
                       "E. Race",
                       "F. Gender",
                       "G. Knowing people",
                       "H. Political connections",
                       "I. Well job",
                       "J. Hard job",
                       "K. Support family",
                       "L. Has children")
colnames(cor_issp) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", 
                       "(H)", "(I)","(J)","(K)","(L)")
corrplot::corrplot(cor_issp,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(12),
  bg = "white",
  na.label = "-",
  lower.col = "black")
```

@@

```{r sum-fit-issp}
load(file = "input/data/proc/data_s1.RData")
pacman::p_load(dplyr, kableExtra, knitr)

table_format = if(knitr::is_html_output()) {
  "html"
} else if(knitr::is_latex_output()) {
  "latex"
}

fsize = if(knitr::is_html_output()) {
  20
} else if(knitr::is_latex_output()) {
  12
}

cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,ordered = T)

cfa_perpref2 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  parents =~ wealthy + pareduc
	  backgrd =~ race + gender
	  networks=~ people + polcone
	  percnmerit=~ parents + backgrd + networks
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = T)

@@

# Create table
table_issp_fits <-
  knitr::kable(sum_fit_issp, format=table_format, digits=3, booktabs=T, 
               col.names=colnames,  
               caption = "Summary fit indices according to model", 
               escape = FALSE) %>%
  kableExtra::kable_styling(font_size = fsize,
                            full_width = F,
                            bootstrap_options=c("striped", "bordered"))
table_issp_fits
```

Table \@ref(tab:sum-fit-issp) shows the four-factor model fit allowing comparison of first and second-order factors. The model restricted to four first-order factor dimensions shows a regular fit (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)) in contrast to the model considering a second-order factor for non-meritocratic perception that has a slightly better fit based on the standard fit measures (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).

@@

```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy",out.width='100%'}
knitr::include_graphics('output/images/cfa_issp_study1.PNG')
```

@@

```{r tb-perpref1, echo=FALSE}
load(file = "input/data/proc/data_s1.RData")
pacman::p_load(dplyr,knitr,kableExtra)
options(knitr.kable.NA = '')
table_format = if(knitr::is_html_output()) {
  "html"
} else if(knitr::is_latex_output()) {
  "latex"
}

@@

cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,
                            ordered = c("wealthy","pareduc","ambition","hwork",
                                        "people","polcone","race","gender",
                                        "respons","yeduc",
                                        "family","child",
                                        "welljob","hardjob"))
labs <- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Reasons for pay: well job",
  "Reasons for pay: hard job",
  "Reasons for pay: support family",
  "Reasons for pay: has children"
)

@@

# Table --------------------------------------------------------------------.
tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA = Root mean square error of approximation")
tb.col <- c("","Meritocratic","Non-meritocratic",
            "Meritocratic","Non-meritocratic")
tb.caption <- c(" Factor loadings and fit measures for ISSP 2009")
tb.issp<- kable(tb.perpref1,escape = FALSE,digits = 3,
                align = "lcccc",
                col.names = tb.col, 
                caption = tb.caption, 
                booktabs = T,
                linesep = "",
                format = table_format) %>%
          kable_styling(latex_options = c("hold_position"),
                        bootstrap_options = c("striped", "bordered"),
                        position = "center",
                        font_size = 12) %>%
          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
          row_spec(row = 12,hline_after = TRUE) %>%
          add_indent(c(13:17)) %>%
          footnote(general =tb.foot ,footnote_as_chunk = T,threeparttable = T);tb.issp
```

### 2021-07-08 processing/preparation_study1.Rmd
ajustes
- cambio a data_s1.RData
- Modificación en data prep de study1. Se detalla mejor el procedimiento.
- check render desde consola  html (OK) y pdf (OK)

@@

```{r}
issp2009 <- read_dta("../input/data/original/ZA5400_v4-0-0.dta")
```

@@

@@

@@

@@

```{r}
merit_per_pref   <- left_join(getahead, detearn, by = c("id", "v5"))

```

# Summary
```{r,results='asis', echo=FALSE}
st_css() # esto configura el css de dfsummary
print(dfSummary(merit_per_pref),
      method = "render")
```

```{r}
save(merit_per_pref, file = "../input/data/proc/merit_per_pref.rda")
```

>>>>

This is the preparation code of the data for Study 1: Measuring meritocracy with international secondary data. The prepared data is `data_s1.RData`

@@

```{r}
issp2009 <- read_dta("input/data/original/ZA5400_v4-0-0.dta")
```

First, we check country identification variables and generate a specific individual-country variable.

@@

The following variables are those who are conceptualized as meritocratic and Non-meritocratic perception.

@@

First, we rename the variables to a more substantial name. In second place, we check for missing values and declare them as NA. Then, we reverse code all the variables, because number 1 represents the "Essential" and 5 represents "Not important at all". Finally, whe adjust variable labels to a more intuitive one.

@@

* v47: responsibility job
* v48: year on education
* v49: support family
* v50: has children to support
* v51: well job
* v52: hard job
* IDs: v5 (country) & id (individual id)

Same as the meritocratic perception. First, we rename the variables to a more substantial name. In second place, we check for missing values and declare them as NA. Then, we reversecode al the variables, because number 1 represents the "Esential" and 5 represents "Not important at all". Finally, whe adjust variable labels to a more intuitive one.

@@

We merge the meritocratic perceptions and preferences variables into a single dataset.

```{r}
data_s1   <- left_join(getahead, detearn, by = c("id", "v5"))
```

# Final summary
```{r,results='asis', echo=FALSE}
st_css() # esto configura el css de dfsummary
print(dfSummary(data_s1),
      method = "render")
```

```{r}
save(data_s1, file = "input/data/proc/data_s1.RData")
```
